---
output: 
  github_document:
    toc: false
    toc_depth: 3

###
### Bibliography settings
###
bibliography: ./inst/REFERENCES.bib
csl:  ./inst/chicago-author-date.csl
suppress-bibliography: false
link-citations: true
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# networkflow

<!-- badges: start -->
[![R-CMD-check](https://github.com/agoutsmedt/networkflow/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/agoutsmedt/networkflow/actions/workflows/R-CMD-check.yaml)
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental)
<!-- badges: end -->

The goal of networkflow (a workflow for networks) is to propose a series of functions to make 
it easier and quicker to manipulats networks. It mainly targets working on bibliometric networks 
(see the [biblionetwork](https://github.com/agoutsmedt/biblionetwork) package for creating such networks). 
This package heavily relies on [igraph](https://igraph.org/r/) and [tidygraph](https://tidygraph.data-imaginist.com/index.html), 
and aims at producing ready-made networks for projecting them using [ggraph](https://ggraph.data-imaginist.com/). 
This package aims at helping the users to follow more quickly and easily the main steps of network manipulation, 
from creating the graph, through detecting clusters, to projecting it. Please see 
`vignette("workflow-network")` for details on the workflow for dealing with a unique
network.

Networkflow also proposes a worfklow to deal with a list of networks, in order to develop a 
dynamic analysis. It implements a method to merge clusters across successive networks,
to identify inter-temporal clusters. It also develops corresponding visualisations to 
display the evolution of clusters across networks. `vignette("exploring_dynamic_networks")`
gives an example of the workflow for dynamic networks. You can also find illustrations for this
method in ["An Independent European Macroeconomics? A History of European Macroeconomics
through the Lens of the European Economic Review](https://aurelien-goutsmedt.com/publication/eer-history/).



You can cite this package as:

```{r}
citation("networkflow")
```


## Installation

You can install the development version from [GitHub](https://github.com/) with:

``` r
install.packages("devtools")
devtools::install_github("agoutsmedt/networkflow")
```

