---
output: 
  github_document:
    toc: false
    toc_depth: 3
bibliography: ./inst/REFERENCES.bib
csl:  ./inst/chicago-author-date.csl
suppress-bibliography: false
link-citations: true
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# networkflow

<!-- badges: start -->
[![R-CMD-check](https://github.com/agoutsmedt/networkflow/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/agoutsmedt/networkflow/actions/workflows/R-CMD-check.yaml)
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental)
<!-- badges: end -->

`networkflow` provides a complete workflow to build, structure, and explore
networks from tabular data.

Its key feature is a built-in dynamic analysis workflow: the package can
build networks across time windows, detect clusters in each window, and link
clusters across periods to track their evolution.

More broadly, `networkflow` supports the full analysis pipeline, from network
construction to interpretation and visualization, including clustering, layout
and color preparation, static plotting, and interactive exploration with a
Shiny app.

The package was developed with projected networks in mind (for example,
article -> reference), but it can also be used more generally once data are
represented as `tbl_graph` objects.

The package includes:

- network construction (`build_network()`, `build_dynamic_networks()`),
- clustering and inter-temporal matching (`add_clusters()`, `merge_dynamic_clusters()`),
- interpretation (`name_clusters()`, `extract_tfidf()`),
- visualization (`layout_networks()`, `color_networks()`, `plot_networks()`),
- interactive exploration (`launch_network_app()`).

For a full walkthrough, see:

- `vignette("networkflow_presentation")`
- https://agoutsmedt.github.io/networkflow/

## Installation

You can install the development version from [GitHub](https://github.com/) with:

``` r
install.packages("devtools")
devtools::install_github("agoutsmedt/networkflow")
```

## Quick start

```{r eval = FALSE}
library(networkflow)

nodes <- subset(Nodes_stagflation, source_type == "Stagflation")
references <- Ref_stagflation

g <- build_network(
  nodes = nodes,
  directed_edges = references,
  source_id = "source_id",
  target_id = "target_id",
  projection_method = "structured",
  cooccurrence_method = "coupling_similarity",
  edges_threshold = 1,
  keep_singleton = FALSE
)

g <- add_clusters(
  graphs = g,
  clustering_method = "leiden",
  objective_function = "modularity",
  seed = 123
)

g <- layout_networks(g, node_id = "source_id", layout = "kk")
g <- color_networks(g, column_to_color = "cluster_leiden")

plot_networks(
  graphs = g,
  x = "x",
  y = "y",
  cluster_label_column = "cluster_leiden",
  node_size_column = NULL,
  color_column = "color"
)
```
