---
output: 
  github_document:
    toc: true
    toc_depth: 3

###
### Bibliography settings
###
bibliography: ./inst/REFERENCES.bib
csl:  ./inst/chicago-author-date.csl
suppress-bibliography: false
link-citations: true
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# networkflow

<!-- badges: start -->
<!-- badges: end -->

The goal of networkflow (a workflow for networks) is to propose a serie of functions to make it easier and quicker to work on networks. It mainly targets working on bibliometric networks (see the [biblionetwork](https://github.com/agoutsmedt/biblionetwork) package for creating such networks). This package heavily relies on [igraph](https://igraph.org/r/) and [tidygraph](https://tidygraph.data-imaginist.com/index.html), and aims at producing ready-made networks for projecting them using [ggraph](https://ggraph.data-imaginist.com/). This package does not invent nothing new, properly speaking, but it allows the users to follow more quickly and easily the main steps of network manipulation, from creating the graph to projecting it. It is inspired by what could be done with [GEPHI](https://gephi.org/): the package allows the use of the Leiden community detection algorithm, as well as of the Force Atlas 2 layout, both being unavailable in igraph (and so in tidygraph).

## Installation

You can install the development version from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("agoutsmedt/networkflow")
```

## The workflow for networks

### First step: creating the network and keeping the main component

As soon as you have a nodes and a edges file (see the [biblionetwork](https://github.com/agoutsmedt/biblionetwork) package for creating such files), you can create a graph, using tidygraph and the [tbl_graph()](https://rdrr.io/cran/tidygraph/man/tbl_graph.html) function. The next step, as it is recurrent in many network analyses, notably in bibliometric netwoks like bibliographic coupling networks would be to keep only the [main component](https://en.wikipedia.org/wiki/Component_(graph_theory)) of your network. This could be done in one step using the `tbl_main_component()` function of `networkflow`. 

```{r creating graph}
library(networkflow)

## basic example code

graph <- tbl_main_component(nodes = Nodes_coupling, edges = Edges_coupling, directed = FALSE, node_key = "ItemID_Ref", nb_components = 1)
print(graph)

```

The parameter `nb_components` allows you to choose the number of components you want to keep. For obvious reasons, it is settled to 1 by default. 

However, it could happen in some networks (for instance co-authorship networks) that the second biggest component of your network is quite large. To avoid removing too big components without knowing it, the `tbl_main_component()` function integrates a warning that happens when a secondary component gathering more than x% of the total number of nodes is removed. The `threshold_alert` parameter is set to 0.05 by default, but you can reduce it if you really want to avoid removing relatively big components.

```{r components}

## basic example code

graph <- tbl_main_component(nodes = Nodes_coupling, edges = Edges_coupling, directed = FALSE, node_key = "ItemID_Ref", threshold_alert = 0.001)
print(graph)

```

### Second step: finding communities

Once you have you tidygraph graph, an important step is to run community detection algorithms to group the nodes depending on their links. This package uses the [leidenAlg](https://cran.r-project.org/web/packages/leidenAlg/index.html) package, and its `find_partition()` function, to implement the Leiden algorithm [@traag2019]. The `leiden_workflow()` function of our package runs the Leiden algorithm and attributes a community number to each node in the `Com_ID` column, but also to each edge (depending if the `from` and `to` nodes are within the same community). 

```{r leiden}

# creating again the graph
graph <- tbl_main_component(nodes = Nodes_coupling, edges = Edges_coupling, directed = FALSE, node_key = "ItemID_Ref", nb_components = 1)

# finding communities
graph <- leiden_workflow(graph)
print(graph)

```

You can observe that the function also gives the size of the community, by calculating the share of total nodes that are in each community.

The function also allows to play with the `resolution` parameter of leidenAlg [`find_partition()`]() function. Varying the resolution of the algorithm results in a different partition and different number of communities. A lower resolution means less communities, and conversely. The basic resolution of the `leiden_workflow()` is set by `res_1` and equals 1 by default. You can vary this parameter, but also try a second resolution with `res_2` and a third one with `res_3`:

```{r resolution}

# creating again the graph
graph <- tbl_main_component(nodes = Nodes_coupling, edges = Edges_coupling, directed = FALSE, node_key = "ItemID_Ref", nb_components = 1)

# finding communities
graph <- leiden_workflow(graph, res_1 = 0.5, res_2 = 2, res_3 = 3)
print(graph)

```

Once you have detected different communities in your network, you are well on the way of the projection of your graph, but two important steps should be implemented before. First, you have to attribute some colors to each community. These colors will be used for your nodes and edges when you will project your graph with `ggraph`. The function `community_colors` of the `networkflow` package allow to do that. You just have to give it a palette (with as many colors as the number of communities for a better visualisation).^[If two connected nodes are in the same community, their edge will take the same color. If they are in different communities, their edge will have a mix of the two communities colors.] 

```{r}
# loading a palette with many colors
palette <- c("#1969B3","#01A5D8","#DA3E61","#3CB95F","#E0AF0C","#E25920","#6C7FC9","#DE9493","#CD242E","#6F4288","#B2EEF8","#7FF6FD","#FDB8D6","#8BF9A9","#FEF34A","#FEC57D","#DAEFFB","#FEE3E1","#FBB2A7","#EFD7F2","#5CAADA","#37D4F5","#F5779B","#62E186","#FBDA28","#FB8F4A","#A4B9EA","#FAC2C0","#EB6466","#AD87BC","#0B3074","#00517C","#871B2A","#1A6029","#7C4B05","#8A260E","#2E3679","#793F3F","#840F14","#401C56","#003C65","#741A09","#602A2A","#34134A","#114A1B","#27DDD1","#27DD8D","#4ADD27","#D3DD27","#DDA427","#DF2935","#DD27BC","#BA27DD","#3227DD","#2761DD","#27DDD1")

# creating again the graph
graph <- tbl_main_component(nodes = Nodes_coupling, edges = Edges_coupling, directed = FALSE, node_key = "ItemID_Ref", nb_components = 1)

# finding communities
graph <- leiden_workflow(graph)

# attributing colors
graph <- community_colors(graph, palette, community_column = "Com_ID")
print(graph)
```

What you want to do next is to give a name automatically to your community. The `community_names()` function allows you to do that: it gives to the community the label of the node, within the community, which has the highest score in the statistics you choose. In the next exemple, we will calculate the degree of each node, and each community will take as a name the label of its highest degree node.

```{r naming}

library(magrittr)
library(dplyr)
library(tidygraph)

# calculating the degree of nodes
 graph <- graph %>%
   activate(nodes) %>%
   mutate(degree = centrality_degree())
 
# giving names to communities
 graph <- community_names(graph, ordering_column = "degree", naming = "Author_date", community_column = "Com_ID")
 print(graph)
 
```


### Third step: giving a structure to your graph

You now have your network with your communities (and names and colors for these communities). The next step is to use layout, that is an algorithm that will attribute coordinates to your network. Here, we use the Force Atlas 2 algorithm [@jacomy2014], which is implemented in the [vite](https://github.com/ParkerICI/vite) package, via the [complete_forceatlas2()](https://rdrr.io/github/ParkerICI/scgraphs/man/complete_forceatlas2.html) function. As it is not on CRAN, you need to install this package via github. 

```{r force_atlas}

if ("vite" %in% installed.packages()==FALSE){devtools::install_github("ParkerICI/vite")}
  library(vite)

graph <- graph %>%
  activate(nodes) %>% 
  mutate(size = degree) # use here what you will use for the size of the nodes in the graph. Force Atlas will take care of avoiding overlapping

graph <- vite::complete_forceatlas2(graph, first.iter = 5000, overlap.method = "repel", overlap.iter = 500)


```

### Fourth step: projecting your graph, at last!

We now have our communities, with their respective colors, and a set of coordinates for each node. Two last steps are necessary to create a more understandable visualisation:

- Selection a list of nodes you want the names to be displayed on the graph (considering the fact that displaying the names of all nodes would produce a too crowded visualisation);
- Giving coordinates to communities labels, for these labels to be displayed on the graph.

The first operation is done using the `top_nodes()` function, which selects a number n of nodes with the highest chosen statistics value per communities and a number n of nodes with the highest chosen statistics value within the whole network. Here, we will use the nodes degree (that we have already calculated above): we will take the node that has the highest degree in each community, as well as the 10 nodes which have the highest degree value in the whole network (of course, the function takes care for doublons).

```{r}
top_nodes  <- top_nodes(graph, ordering_column = "degree", top_n = 10, top_n_per_com = 1)
print(top_nodes)
```


The second operation is done using the `community_labels()` function, which calculate the mean of coordinates x and y for each community. These coordinates are to be used to plot the name of the community on the graph visualisation. 

```{r}
community_labels <- community_labels(graph, community_name_column = "Community_name", community_size_column = "Size_com")
print(labels)
```


We can now plot the graph using the [ggraph](https://ggraph.data-imaginist.com/) package!

```{r fig.width = 12, fig.height = 10}
library(ggraph)
library(ggrepel) # use for avoiding labels overlapping
library(ggnewscale) # this is needeed to use several size scale in the graph (here, one for nodes, and one for labels)

ggraph(graph, "manual", x = x, y = y) + 
  geom_edge_arc0(aes(color = color_edges, width = weight), alpha = 0.4, strength = 0.2, show.legend = FALSE) +
  scale_edge_width_continuous(range = c(0.1,2)) +
  scale_edge_colour_identity() +
  geom_node_point(aes(x=x, y=y, size = degree, fill = color), pch = 21, alpha = 0.9, show.legend = FALSE) +
  scale_size_continuous(range = c(0.2,13)) +
  scale_fill_identity() +
  new_scale("size") +
  geom_text_repel(data=top_nodes, aes(x=x, y=y, label = Label), size = 2, fontface="bold", alpha = 1, point.padding=NA, show.legend = FALSE) +
  geom_label_repel(data=community_labels, aes(x=x, y=y, label = Community_name, fill = color), size = 6, fontface="bold", alpha = 0.9, point.padding=NA, show.legend = FALSE) +
  scale_size_continuous(range = c(0.5,5)) +
  theme_void()
```


## References
