---
title: "Presentation of networkflow"
author: ""
description: ""
output:
  rmarkdown::html_vignette:
    toc: true
vignette: >
  %\VignetteIndexEntry{Presentation of networkflow}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

# Introduction 

## Overview

`networkflow` provides a complete workflow to build, structure, and explore networks from tabular data.

Its key feature is a built-in dynamic analysis workflow: the package can build networks across time windows, detect clusters in each window, and link clusters across periods to track their evolution.

More broadly, `networkflow` supports the full analysis pipeline, from network construction to interpretation and visualization, including clustering, layout and color preparation, static plotting, and interactive exploration with a Shiny app.

The package was developed with projected networks in mind (for example, article -> reference), but it can also be used more generally once data are represented as `tbl_graph` objects.

## What this package does

The package is organized in three main steps:

1. create networks (static or dynamic) from tabular data;
2. detect and harmonize clusters across time windows;
3. prepare visualization and exploration outputs (layout, colors, labels, plotting, Shiny app).


## Typical workflow

A typical workflow is:

1. prepare `nodes` and `directed_edges` tables;
2. build the network with `build_network()` (or `build_dynamic_networks()` for temporal analyses);
3. detect clusters with `add_clusters()`;
4. prepare plotting attributes (`layout_networks()`, `color_networks()`);
5. inspect and interpret results with plots and `launch_network_app()`.

# Data Requirements

## Network objects: `tbl_graph`

A **`tbl_graph`** is the core network object in `networkflow`.

It comes from `tidygraph` and stores:

1. a node table (attributes of entities);
2. an edge table (connections between entities).

Most functions in this package take a `tbl_graph` (or a list of `tbl_graph`) as input. In practice, `build_network()` and `build_dynamic_networks()` are the main entry points that create these objects from tabular data.

`networkflow` expects tabular inputs with explicit identifiers:

1. `nodes`: one row per source entity (for example, one row per article), with a unique ID used as `source_id`;
2. `directed_edges`: links from `source_id` to `target_id` (for example, article -> reference, author -> paper);
3. `time_variable`: required only for dynamic analyses with `build_dynamic_networks()`.


## Input of `build_network()` and `build_dynamic_networks()`

`build_network()` and `build_dynamic_networks()` are the main entry points to create `tbl_graph` objects in the package. They start from a bipartite relation (`source_id` -> `target_id`) and produce a one-mode weighted network on `source_id` entities. The bipartite relation is a table of directed edges from source to target entities (for example, article -> reference, author -> paper). After projection, the result is a one-mode network.

If your data is already one-mode (for example, author -> author), you can build a `tbl_graph` directly and use downstream functions for clustering, layout, plotting, and exploration. Typical downstream functions in this case are `add_clusters()`, `layout_networks()`, `color_networks()`, and `launch_network_app()`.

# Step 1: Creating networks

Functions used:

- `build_dynamic_networks()`
- `build_network()`
- `filter_components()`

This step creates one static network (`tbl_graph`) or a list of temporal networks
(`list` of `tbl_graph`) from bipartite links (`source_id -> target_id`).
`build_network()` is the single-network wrapper around `build_dynamic_networks()`.

### `build_network()` and `build_dynamic_networks()`


`build_dynamic_networks()` builds one network or a list of time-window networks if the user provides a temporal variable. `build_network()` is the wrapper for one network of `build_dynamic_networks(time_variable = NULL)`. `build_dynamic_networks()` supports two different filtering strategies:

`build_dynamic_networks()` takes as input a bipartite relation and projects it into a one-mode network. It supports two different filtering strategies for edge retention after projection:

1. a structured strategy, which defines edge strength using measures derived from co-occurrence intensity;
2. a statistical strategy, which defines a null model of random co-occurrence and keeps edges based on statistical significance. 

In short, the first approach defines and filters by observed tie strength, while the second defines and filters by statistical significance. The structured method is generally more computationally efficient, while the statistical method provides a more rigorous filter for connections beyond random chance. For example, if `source_id` is article and `target_id` is reference, the structured method will keep article pairs with strong co-citation or bibliographic coupling, while the statistical method will keep article pairs whose co-citation or bibliographic coupling is significantly stronger than expected under a random model.

Parameters common to both methods:

- `nodes`: table with one row per source entity defined by `source_id`.
- `directed_edges`: table with directed edges from `source_id` to `target_id`.
- `source_id`, `target_id`: identifier columns used for projection.
- `projection_method`: `"structured"` or `"statistical"`.
- `compute_size`: if `TRUE`, computes `node_size`.
- `keep_singleton`: if `FALSE`, removes isolated nodes.

Structured-only parameters (`projection_method = "structured"`)

- uses cooccurrence/coupling measures from the `biblionetwork` package.
- `cooccurrence_method`: `"coupling_angle"`, `"coupling_strength"`, `"coupling_similarity"`.
- `edges_threshold`: minimum edge strength retained.


Statistical-only parameters (`projection_method = "statistical"`)

- uses statistical backbone extraction (`backbone` package).
- `model`: `"sdsm"`, `"fdsm"`, `"fixedfill"`, `"fixedrow"`, `"fixedcol"`.
- `alpha`: significance threshold for edge retention.
- `backbone_args`: additional arguments passed to backbone routines.

Dynamic-only parameters (`build_dynamic_networks()`):

- `time_variable`: temporal column in `nodes` (for example publication year).
- `time_window`: width of each window.
- `overlapping_window`: rolling windows (`TRUE`) or disjoint windows (`FALSE`). In the first case, partition is done by rolling the time window by one unit (for example, 1990-2009, 1991-2010, etc.). In the second case, partition is done by fixed intervals (for example, 1990-1999, 2000-2009, etc.).

The main output of these functions is a `tbl_graph` (or a list of `tbl_graph` for dynamic analyses). If `projection_method = "structured"`, the output edges are weighted by the selected cooccurrence/coupling measure. If `projection_method = "statistical"`, the output edges are unweighted and retained based on statistical significance.

Examples: 

```{r static-build-setup}
library(networkflow)

nodes <- subset(Nodes_stagflation, source_type == "Stagflation")

references <- Ref_stagflation
```

```{r static-build-network}
g_static <- build_network(
  nodes = nodes,
  directed_edges = references,
  source_id = "source_id",
  target_id = "target_id",
  projection_method = "structured",
  cooccurrence_method = "coupling_similarity",
  edges_threshold = 1,
  compute_size = FALSE,
  keep_singleton = FALSE
)
```

```{r dynamic-build-network-structured, eval = FALSE}
g_dynamic <- build_dynamic_networks(
  nodes = nodes,
  directed_edges = references,
  source_id = "source_id",
  target_id = "target_id",
  time_variable = "source_year",
  time_window = 20,
  projection_method = "structured",
  cooccurrence_method = "coupling_similarity",
  edges_threshold = 1,
  overlapping_window = TRUE,
  compute_size = FALSE,
  keep_singleton = FALSE
)
```

```{r dynamic-build-network-statistical, eval = FALSE}
g_dynamic_stat <- build_dynamic_networks(
  nodes = nodes,
  directed_edges = references,
  source_id = "source_id",
  target_id = "target_id",
  time_variable = "source_year",
  time_window = 20,
  projection_method = "statistical",
  model = "sdsm",
  alpha = 0.05,
  overlapping_window = TRUE,
  compute_size = FALSE,
  keep_singleton = FALSE
)
```

### `filter_components()` 

Use `filter_components()` to keep the main connected component(s):

- `nb_components`: number of largest components to keep.
- `threshold_alert`: warning threshold when a removed component is still large.
- `keep_component_columns`: keep or remove helper columns on component IDs and sizes.

```{r static-filter-components, eval = FALSE}
g_static <- filter_components(g_static, nb_components = 1)
```


# Step 2: Clustering

Functions used:

- `add_clusters()`
- `merge_dynamic_clusters()`
- `name_clusters()`
- `add_node_roles()`
- `extract_tfidf()`

### `add_clusters()`

Run community detection on a static or dynamic network. The function is a wrapper around `tidygraph::group_graph()`. It also supports the `igraph` implementation of the Leiden algorithm, which is the default method. 

Main parameters:

- `clustering_method`: the clustering algorithm to use. 
- `weights`: edge weight column usage.
- `objective_function`, `resolution`, `n_iterations`: Leiden controls.
- `seed`: reproducibility for stochastic algorithms.

The output is a `tbl_graph` (or a list of `tbl_graph`) with new columns: 
- node column `cluster_{method}`.
- edge columns `cluster_{method}_from`, `cluster_{method}_to`, `cluster_{method}`.
- node column `size_cluster_{method}` with cluster shares.

Example: 

```{r static-add-clusters-leiden}
g_static <- add_clusters(
  graphs = g_static,
  clustering_method = "leiden",
  objective_function = "modularity",
  resolution = 1,
  n_iterations = 1000,
  seed = 123
)
```

```{r dynamic-add-clusters-leiden, eval = FALSE}
g_dynamic <- add_clusters(
  graphs = g_dynamic,
  clustering_method = "leiden",
  objective_function = "modularity",
  resolution = 1,
  n_iterations = 1000,
  seed = 123
)
```

### `merge_dynamic_clusters()`(dynamic only)

`add_clusters()` runs independently on each time window, so cluster IDs are not directly comparable across windows. `merge_dynamic_clusters()` links clusters from adjacent windows when node overlap is high enough, and assigns stable intertemporal IDs.

Input requirements:

- `list_graph` must be a list of at least two `tbl_graph`.
- the list order must be chronological (oldest to most recent window).

Main parameters:

- `cluster_id`: input cluster column (for example `cluster_leiden`).
- `node_id`: stable node identifier across windows.
- `threshold_similarity`: matching threshold in `(0.5, 1]`.
- `similarity_type`: `"complete"` or `"partial"`.

`similarity_type` controls how overlap is computed:

- `"complete"`: the overlap share is computed over all nodes in the compared clusters, including entries that exist only in one window. This is stricter when network size changes over time.
- `"partial"`: the overlap share is computed only on nodes present in both adjacent windows. This is often preferable when many new nodes enter over time.

Output:

- new node column `dynamic_{cluster_id}` (for example `dynamic_cluster_leiden`). The dynamic cluster IDs are assigned by propagation: it starts by assigning unique IDs to clusters in the first time window, then propagates those IDs to later windows when a cluster match passes the similarity threshold; otherwise, a new dynamic ID is created.
- corresponding edge columns `dynamic_{cluster_id}_from`, `dynamic_{cluster_id}_to`, and `dynamic_{cluster_id}`.

```{r dynamic-merge-clusters, eval = FALSE}
g_dynamic <- merge_dynamic_clusters(
  list_graph = g_dynamic,
  cluster_id = "cluster_leiden",
  node_id = "source_id",
  threshold_similarity = 0.51,
  similarity_type = "partial"
)
```

### `name_clusters()`

Cluster IDs are not very informative in themselves. `name_clusters()` helps assign readable labels to clusters based on their content. The labels are not meant to be definitive cluster names, but rather a quick way to get a sense of cluster content. The function supports three methods:

- `method = "tf-idf"`: labels clusters with the most distinctive terms extracted from a `text_columns`. This is usually the best default for thematic interpretation.
- `method = "given_column"`: selects, within each cluster, the node with the
  highest value in `order_by`, then builds the label from `label_columns` of
  that node. Typically, you can use this method to label clusters with the title of a representative article (for example the most cited one).
- `method = "tidygraph_functions"`: computes a centrality measure with
  `tidygraph_function`, selects the most central node per cluster, then builds
  the label from `label_columns`.

Main parameters:

- `method`: `"tidygraph_functions"`, `"given_column"`, or `"tf-idf"`.
- `name_merged_clusters`: `TRUE` to name dynamic clusters across the list. Typically, you want to set this to `TRUE` when your `cluster_id` is the dynamic cluster column created by `merge_dynamic_clusters()`.
- `cluster_id`: column to name.
- `label_name`: output label column name (`"cluster_label"` by default).
- `text_columns`, `nb_terms_label`: key arguments for TF-IDF naming.

```{r dynamic-name-clusters, eval = FALSE}
g_dynamic <- name_clusters(
  graphs = g_dynamic,
  method = "tf-idf",
  name_merged_clusters = TRUE,
  cluster_id = "dynamic_cluster_leiden",
  text_columns = "source_title",
  nb_terms_label = 3
)
```

### `add_node_roles()`

Nodes in a cluster can play different structural roles. `add_node_roles()` implements the Guimera-Amaral classification of node roles based on two measures: within-module degree (z-score) and participation coefficient. Use `add_node_roles()` after clustering to classify nodes according to their structural position in modules (within-module degree, participation coefficient, Guimera-Amaral roles). This helps distinguish peripheral nodes, connectors, and hubs.


Main parameters:

- `module_col`: cluster/module column used to compute roles.
- `weight_col`: edge weight column.
- `z_threshold`: hub threshold for within-module z-score.

Main outputs:

- `within_module_degree`.
- `within_module_z`.
- `participation_coeff`.
- `role_ga`.

```{r static-node-roles, eval = FALSE}
g_static <- add_node_roles(
  graphs = g_static,
  module_col = "cluster_leiden",
  weight_col = "weight",
  z_threshold = 2.5
)
```

### `extract_tfidf()`

Use `extract_tfidf()` to characterize cluster content from textual metadata (for instance titles, abstracts, or keywords). In a static network, cluster IDs are usually unique within the graph, so `grouping_across_list = FALSE`.
Main parameters:

- `text_columns`: one or more text fields used to extract ngrams.
- `grouping_columns`: document units for TF-IDF (for example cluster IDs).
- `grouping_across_list`: helps disambiguate group IDs across windows.
- `n_gram`: maximum n for ngrams.
- `clean_word_method`: `"lemmatize"`, `"stemming"`, `"none"`.
- `ngrams_filter`: remove terms that are too rare globally.
- `nb_terms`: number of top terms returned per group.

```{r static-tfidf, eval = FALSE}
tfidf_static <- extract_tfidf(
  data = g_static,
  text_columns = "source_title",
  grouping_columns = "cluster_leiden",
  grouping_across_list = FALSE,
  n_gram = 2,
  nb_terms = 5
)
```

# Step 3: Plot networks

Functions used:

- `layout_networks()`
- `minimize_crossing_alluvial()`
- `color_networks()`
- `prepare_label_alluvial()`
- `prepare_label_networks()`
- `plot_alluvial()`
- `plot_networks()`
- `launch_network_app()`

### `layout_networks()`

Compute node coordinates before plotting. The function is a wrapper around `ggraph::ggraph()` and supports all its layout algorithms. For dynamic networks, coordinates are computed sequentially by window: the first window is computed with the selected layout, then subsequent windows are computed by reusing prior coordinates when `compute_dynamic_coordinates = TRUE`.

Main parameters:

- `node_id`: unique node ID column used to join coordinates.
- `layout`: layout algorithm accepted by `ggraph::create_layout()`.
- `compute_dynamic_coordinates`: reuse prior window coordinates.
- `save_coordinates`: if `TRUE`, saves coordinates in node columns `{layout}_x` and `{layout}_y` (for example `kk_x`, `kk_y`). Typically, you want to set this to `TRUE` when testing different layouts for plotting. 

The output is a `tbl_graph` (or list of `tbl_graph`) with new node columns `{layout}_x` and `{layout}_y` or `x` and `y` if `save_coordinates = FALSE`. 

Example: 

```{r static-layout}
g_static <- layout_networks(
  graphs = g_static,
  node_id = "source_id",
  layout = "kk"
)
```

```{r dynamic-layout, eval = FALSE}
g_dynamic <- layout_networks(
  graphs = g_dynamic,
  node_id = "source_id",
  layout = "fr",
  compute_dynamic_coordinates = TRUE
)
```

### `color_networks()`

Assign colors to nodes and edges based on a categorical attribute `column_to_color` present in the node table. Typically, it is used to color clusters from `add_clusters()`. The function supports various color input formats: a named vector of colors with a length equal to the number of unique categories in `column_to_color`, a data frame mapping categories to colors. If `color = NULL`, the function generates a color palette automatically. 

Main parameters:

- `column_to_color`: node attribute used to define categories to color.
- `color`: : a palette or a two-column data frame mapping categories to colors. 
- `unique_color_across_list`: for dynamic networks only. It controls whether the same value of `column_to_color` in different time windows should receive the same color. If set to `FALSE`, the same categorical variable will be considered as the same variable in different graphs. If set to `TRUE`, the same categorical variable will be considered as a different variable in different graphs and thus receive a different color.

Output:

- node column `color`.
- edge column `color` computed as a mix of source and target node colors.

```{r static-colors}
g_static <- color_networks(
  graphs = g_static,
  column_to_color = "cluster_leiden",
  color = NULL
)
```

### `prepare_label_networks()`

Create label coordinates (`label_x`, `label_y`) for the label positioning in network plots. The function computes the average coordinates of nodes within each cluster to position the label. 

Main parameters:

- `x`, `y`: coordinate columns used to compute label centers.
- `cluster_label_column`: column used for grouping and label text.

```{r static-labels}
g_static <- prepare_label_networks(
  graphs = g_static,
  x = "x",
  y = "y",
  cluster_label_column = "cluster_leiden"
)
```

The output is a `tbl_graph` (or list of `tbl_graph`) with new node columns `label_x` and `label_y` for label coordinates. 

### `plot_networks()`

`plot_networks()` builds a ready-to-use network visualization from graph attributes (coordinates, colors, labels). For exploration and analysis, we strongly encourage to use launch_network_app() for interactive exploration. 

It requires node coordinates (x, y) and a cluster label column. Colors must either already exist or be generated by setting `color_networks = TRUE`. The user can also customize the plot by setting `print_plot_code = TRUE`, which prints the generated ggplot/ggraph code for manual adjustments. 

Main parameters:

- `x`, `y`: node coordinates.
- `cluster_label_column`: displayed cluster labels.
- `node_size_column`: node size variable (`NULL` or missing column gives constant size).
- `color_column`: color column for nodes and edges.
- `color_networks`: if `TRUE`, applies `color_networks()` automatically with
  `cluster_label_column` as grouping variable.
- `color`: optional palette passed to `color_networks()` when
  `color_networks = TRUE`.
- `print_plot_code`: if `TRUE`, prints the generated ggplot/ggraph code for
  manual customization.

Automatic behavior:

- if label coordinates are missing, `prepare_label_networks()` is called
  automatically.
- if edge weights are missing, a constant weight of `1` is used.
- if `node_size_column` is missing, a constant node size of `1` is used.

The output is a ggplot object. For dynamic analyses, the function returns a list of ggplot objects (one per time window) stored in the `$plot` column of each list element.

```{r static-plot, eval = FALSE}
plot_networks(
  graphs = g_static,
  x = "x",
  y = "y",
  cluster_label_column = "cluster_leiden",
  node_size_column = NULL,
  color_column = "color"
)
```

### `launch_network_app()`

`launch_network_app()` extends `plot_networks()` by providing an interactive
Shiny interface for network exploration. It launches a local app with an
interactive network view. Users can click on clusters to display a table with
selected metadata and adjust visual settings (node size, edge width, labels,
edge visibility) to improve readability. For example, for a coupling network, the application allows users to explore article-level information in each cluster. 

The app expects a `tbl_graph` (or a list of `tbl_graph` for dynamic analysis),
cluster identifiers, and metadata columns to display. If the input is a list,
the app shows a dropdown menu to select the graph by list name (typically time
windows when graphs are built with `build_dynamic_networks()`).

Main parameters:

- `cluster_id`: node cluster column used for interaction.
- `cluster_information`: node metadata columns shown in the table (for example `c("source_author", "source_title", "source_year")`), present in node data.
- `node_id`: unique node ID.
- `node_tooltip`: optional node tooltip column for hover information.
- `node_size`: optional node size column.
- `color`: optional color column (`color_networks()` is applied if `NULL`).
- `layout`: layout algorithm available in `layout_networks()`. If `NULL`, the function assumes layout coordinates already exist in node columns `x` and `y`.

```{r static-app, eval = FALSE}
launch_network_app(
  graph_tbl = g_static,
  cluster_id = "cluster_leiden",
  cluster_information = c("source_author", "source_title", "source_year", "source_journal"),
  node_id = "source_id",
  node_tooltip = "source_label",
  node_size = NULL,
  color = "color",
  layout = NULL
)
```

### `prepare_label_alluvial()`

###  `minimize_crossing_alluvial()`

### `plot_alluvial()`



# End-to-end executable example 

The chunk below runs a complete static workflow on bundled data:
network construction, clustering, layout, coloring, label preparation, and plotting.

```{r static-end-to-end, message = FALSE, warning = FALSE}
set.seed(123)

# 1) Input tables
nodes_ex <- subset(Nodes_stagflation, source_type == "Stagflation")

references_ex <- Ref_stagflation

# 2) Build network
g_pipeline <- build_network(
  nodes = nodes_ex,
  directed_edges = references_ex,
  source_id = "source_id",
  target_id = "target_id",
  projection_method = "structured",
  cooccurrence_method = "coupling_similarity",
  edges_threshold = 1,
  keep_singleton = FALSE
)

# 3) Cluster
g_pipeline <- add_clusters(
  graphs = g_pipeline,
  clustering_method = "leiden",
  objective_function = "modularity",
  resolution = 1,
  n_iterations = 1000,
  seed = 123
)

# 4) Prepare plot attributes
g_pipeline <- layout_networks(g_pipeline, node_id = "source_id", layout = "kk")
g_pipeline <- color_networks(g_pipeline, column_to_color = "cluster_leiden")
g_pipeline <- prepare_label_networks(
  g_pipeline,
  x = "x",
  y = "y",
  cluster_label_column = "cluster_leiden"
)

# 5) Quick checks on generated attributes
head(
  g_pipeline %>%
    tidygraph::activate(nodes) %>%
    as.data.frame() %>%
    subset(select = c(source_id, cluster_leiden, size_cluster_leiden, x, y, color, label_x, label_y))
)

head(
  g_pipeline %>%
    tidygraph::activate(edges) %>%
    as.data.frame() %>%
    subset(select = c(from, to, weight, color))
)
```

```{r static-end-to-end-plot, fig.width = 8, fig.height = 6, fig.alt = "Static network plot showing clustered nodes colored by community with labels positioned near cluster centers."}
plot_networks(
  graphs = g_pipeline,
  x = "x",
  y = "y",
  cluster_label_column = "cluster_leiden",
  node_size_column = NULL,
  color_column = "color"
)
```


# Included datasets

- `Nodes_stagflation`
- `Ref_stagflation`
- `Authors_stagflation`
- `Nodes_coupling`
- `Edges_coupling`

# Deprecated functions

- `community_names()` -> use `name_clusters()`
- `community_labels()` -> use `prepare_label_networks()`
- `community_colors()` -> use `color_networks()`
- `dynamic_network_cooccurrence()` -> use `build_dynamic_networks()`
- `intertemporal_cluster_naming()` -> use `merge_dynamic_clusters()`
- `leiden_workflow()` -> use `add_clusters()`
- `networks_to_alluv()`
- `tbl_main_component()` -> use `filter_components()`
- `top_nodes()`
