---
title: "A Workflow for network analysis"
author: "Aurélien Goutsmedt and Alexandre Truc"
description: >
  Introduction to the standards function of the networkflow package for building, manipulating
  plotting, and analysing a network.
output: 
  rmarkdown::html_vignette:
    toc: true
vignette: >
  %\VignetteIndexEntry{workflow-network}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
  

### Bibliography settings
bibliography: REFERENCES.bib
csl:  chicago-author-date.csl
suppress-bibliography: false
link-citations: true
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

In the following article, we will go other the different steps to create a bibliometric network, manipulate it, prepare the plotting and eventually plot it. 

## First step: creating the network and keeping the main component

As a point of departure, you need your bibliometric data to be prepared in a certain format:^[See how to extract and clean data from scopus [here](https://aurelien-goutsmedt.com/post/extracting-biblio-data-1/) and from Dimensions [here](https://aurelien-goutsmedt.com/post/extracting-biblio-data-2/).]

- you need a `nodes` table. For instance, it may be a list of articles with metadata (author(s), title, journal, etc.). Nodes must have a unique identifier and all the information about a node are gathered on only one row (in case of articles, you need one row per article).
- you need a `directed_edges` table, that is a table that links your nodes with another variable that will be used to build the edges between your nodes. For instance, the table could links articles (your nodes) with the references cited by these articles. It can also be a journal or a list of authors (if you are interested in collaboration). In your `directed_edges` table, you need the identifier of the nodes (also present in the `nodes` table), and the unique identifier of the categories (references cited, journals, authors...) the nodes are linked to.   

As soon as you have a nodes and a edges file (see the [biblionetwork](https://github.com/agoutsmedt/biblionetwork) package for creating such files), you can create a graph, using tidygraph and the [tbl_graph()](https://rdrr.io/cran/tidygraph/man/tbl_graph.html) function. The next step, as it is recurrent in many network analyses, notably in bibliometric netwoks like bibliographic coupling networks would be to keep only the [main component](https://en.wikipedia.org/wiki/Component_(graph_theory)) of your network. This could be done in one step using the `tbl_main_component()` function of `networkflow`. 

```{r creating graph}
library(networkflow)

## basic example code

graph <- tbl_main_component(nodes = Nodes_coupling, edges = Edges_coupling, directed = FALSE, node_key = "ItemID_Ref", nb_components = 1)
print(graph)

```

The parameter `nb_components` allows you to choose the number of components you want to keep. For obvious reasons, it is settled to 1 by default. 

However, it could happen in some networks (for instance co-authorship networks) that the second biggest component of your network is quite large. To avoid removing too big components without knowing it, the `tbl_main_component()` function integrates a warning that happens when a secondary component gathering more than x% of the total number of nodes is removed. The `threshold_alert` parameter is set to 0.05 by default, but you can reduce it if you really want to avoid removing relatively big components.

```{r components}

## basic example code

graph <- tbl_main_component(nodes = Nodes_coupling, edges = Edges_coupling, directed = FALSE, node_key = "ItemID_Ref", threshold_alert = 0.001)
print(graph)

```

## Second step: finding communities

Once you have you tidygraph graph, an important step is to run community detection algorithms to group the nodes depending on their links. This package uses the [leidenAlg](https://cran.r-project.org/web/packages/leidenAlg/index.html) package, and its `find_partition()` function, to implement the Leiden algorithm [@traag2019]. The `leiden_workflow()` function of our package runs the Leiden algorithm and attributes a community number to each node in the `Com_ID` column, but also to each edge (depending if the `from` and `to` nodes are within the same community). 

```{r leiden, eval = FALSE}

# creating again the graph
graph <- tbl_main_component(nodes = Nodes_coupling, edges = Edges_coupling, directed = FALSE, node_key = "ItemID_Ref", nb_components = 1)

# finding communities
graph <- leiden_workflow(graph)
print(graph)

```

You can observe that the function also gives the size of the community, by calculating the share of total nodes that are in each community.

The function also allows to play with the `resolution` parameter of leidenAlg [`find_partition()`]() function. Varying the resolution of the algorithm results in a different partition and different number of communities. A lower resolution means less communities, and conversely. The basic resolution of the `leiden_workflow()` is set by `res_1` and equals 1 by default. You can vary this parameter, but also try a second resolution with `res_2` and a third one with `res_3`:

```{r resolution, eval = FALSE}

# creating again the graph
graph <- tbl_main_component(nodes = Nodes_coupling, edges = Edges_coupling, directed = FALSE, node_key = "ItemID_Ref", nb_components = 1)

# finding communities
graph <- leiden_workflow(graph, res_1 = 0.5, res_2 = 2, res_3 = 3)
print(graph)

```

Once you have detected different communities in your network, you are well on the way of the projection of your graph, but two important steps should be implemented before. First, you have to attribute some colors to each community. These colors will be used for your nodes and edges when you will project your graph with `ggraph`. The function `community_colors` of the `networkflow` package allow to do that. You just have to give it a palette (with as many colors as the number of communities for a better visualisation).^[If two connected nodes are in the same community, their edge will take the same color. If they are in different communities, their edge will have a mix of the two communities colors.] 

```{r, eval = FALSE}
# loading a palette with many colors
palette <- c("#1969B3","#01A5D8","#DA3E61","#3CB95F","#E0AF0C","#E25920","#6C7FC9","#DE9493","#CD242E","#6F4288","#B2EEF8","#7FF6FD","#FDB8D6","#8BF9A9","#FEF34A","#FEC57D","#DAEFFB","#FEE3E1","#FBB2A7","#EFD7F2","#5CAADA","#37D4F5","#F5779B","#62E186","#FBDA28","#FB8F4A","#A4B9EA","#FAC2C0","#EB6466","#AD87BC","#0B3074","#00517C","#871B2A","#1A6029","#7C4B05","#8A260E","#2E3679","#793F3F","#840F14","#401C56","#003C65","#741A09","#602A2A","#34134A","#114A1B","#27DDD1","#27DD8D","#4ADD27","#D3DD27","#DDA427","#DF2935","#DD27BC","#BA27DD","#3227DD","#2761DD","#27DDD1")

# creating again the graph
graph <- tbl_main_component(nodes = Nodes_coupling, edges = Edges_coupling, directed = FALSE, node_key = "ItemID_Ref", nb_components = 1)

# finding communities
graph <- leiden_workflow(graph)

# attributing colors
graph <- community_colors(graph, palette, community_column = "Com_ID")
print(graph)
```

What you want to do next is to give a name automatically to your community. The `community_names()` function allows you to do that: it gives to the community the label of the node, within the community, which has the highest score in the statistics you choose. In the next exemple, we will calculate the degree of each node, and each community will take as a name the label of its highest degree node.

```{r naming, eval = FALSE}

library(magrittr)
library(dplyr)
library(tidygraph)

# calculating the degree of nodes
 graph <- graph %>%
   activate(nodes) %>%
   mutate(degree = centrality_degree())
 
# giving names to communities
 graph <- community_names(graph, ordering_column = "degree", naming = "Author_date", community_column = "Com_ID")
 print(graph)
 
```


## Third step: plotting the network

### Preparing the plot



## References
